%%%% ijcai22.tex

%\typeout{IJCAI--22 Instructions for Authors}

% These are the instructions for authors for IJCAI-22.

\documentclass{article}
\pdfpagewidth=8.5in
\pdfpageheight=11in
% The file ijcai22.sty is NOT the same as previous years'
\usepackage{ijcai22}

% Use the postscript times font!
\usepackage{times}
\usepackage{soul}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[small]{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algorithmic}
\urlstyle{same}


% customly used packages
% \usepackage{mathfont}
\usepackage{amssymb}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows, backgrounds, scopes}
\usepackage{pgfplots}
\pgfplotsset{width=6.75cm, compat=newest}
\usepackage[utf8]{inputenc}
\DeclareUnicodeCharacter{2212}{−}
\usepgfplotslibrary{groupplots,dateplot}
\usetikzlibrary{patterns,shapes.arrows}

% the following package is optional:
%\usepackage{latexsym}
\newcommand\todo[1]{\textcolor{red}{Givasile and/or Diou:} \textcolor{red}{#1}}
\newcommand\tododiou[1]{\textcolor{blue}{Diou:} \textcolor{blue}{#1}}
\newcommand\todogiv[1]{\textcolor{purple}{givasile:} \textcolor{purple}{#1}}

\newcommand{\xc}{\mathbf{x}_c}
\newcommand{\Xc}{\mathbf{\mathcal{X}}_c}
\newcommand{\xci}{\mathbf{x}^i_{\mathbf{c}}}
\newcommand{\xb}{\mathbf{x}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Jac}{\mathbf{J}}
% \newcommand{\ale}{f_{ALE}}
% \newcommand{\alea}{\hat{f}_{ALE}}
% \newcommand{\myale}{\tilde{f}_{ALE}}

\usepackage{xfrac}
\usepackage{mleftright}
\usepackage{xparse}
\NewDocumentCommand{\evalat}{sO{\big}mm}{%
  \IfBooleanTF{#1}
   {\mleft. #3 \mright|_{#4}}
   {#3#2|_{#4}}%
}
% See https://www.overleaf.com/learn/latex/theorems_and_proofs
% for a nice explanation of how to define new theorems, but keep
% in mind that the amsthm package is already included in this
% template and that you must *not* alter the styling.
\newtheorem{example}{Example}
\newtheorem{theorem}{Theorem}

% Following comment is from ijcai97-submit.tex:
% The preparation of these files was supported by Schlumberger Palo Alto
% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
% preparation.

% These instructions can be modified and used in other conferences as long
% as credit to the authors and supporting agencies is retained, this notice
% is not changed, and further modification or reuse is not restricted.
% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
% contacts for providing assistance without their prior permission.

% To use for other conferences, change references to files and the
% conference appropriate and use other authors, contacts, publishers, and
% organizations.
% Also change the deadline and address for returning papers and the length and
% page charge instructions.
% Put where the files are available in the appropriate places.

% PDF Info Is REQUIRED.
% Please **do not** include Title and Author information
\pdfinfo{
/TemplateVersion (IJCAI.2022.0)
}

\title{DALE: An efficient and accurate approximation method for ALE plots}

% Multiple author syntax (remove the single-author syntax above and the \iffalse ... \fi here)
% Check the ijcai22-multiauthor.tex file for detailed instructions
%% \author{
%% Vasileios Gkolemis\(^1\)
%% \and
%% Christos Diou\(^2\)\and
%% Theodore Dalamagas \(^{1}\)
%% \affiliations
%% \(^1\)First Affiliation\\
%% \(^2\)Department of Informatics and Telematics, Harokopio University of Athens, Greece
%% \emails
%% \{first, second\}@example.com,
%% cdiou@hua.gr,
%% }

\begin{document}

\maketitle

\begin{abstract}
  Aggregated Local Effect (ALE) is a method for accurate estimation
  and visualization of feature effects, which overcomes fundamental
  failure modes of previously-existed methods, such as Partial
  Dependence Plots (PDP). The \textit{approximation} method of ALE,
  however, faces two crucial weaknesses. Firstly, it does not scale
  well in cases where the input has high dimensionality. Secondly, it
  is vulnerable to out-of-distribution sampling in cases with limited
  training examples. In this paper we propose a novel ALE
  approximation, called Differential Aggregated Local Effects (DALE),
  which overcomes these issues. Our proposal has significant
  computational advantages; Specifically, the computational complexity
  for computing the feature effect for all attributes using DALE is
  similar to what ALE requires for a single attribute. DALE,
  therefore, makes feature effect applicable to high-dimensional
  Machine Learning scenarios with near-zero computational
  overhead. Furthermore, unlike ALE, it calculates the feature effect
  using only the available samples, without creating artificial ones;
  This strategy resolves the weakness of misleading estimations due to
  out-of-distribution sampling. Experiments using both synthetic and
  real datasets demonstrate the value of the proposed approach. Code
  to reproduce these experiments will become publicly available upon
  acceptance.
\end{abstract}

\section{Introduction}
\input{./chapters/1-introduction.tex}

\section{Related Work}
\input{./chapters/2-related.tex}

\section{Method}

In this section, we present the DALE approximation;
In~\ref{sec:3-1-DALE}, we formulate the expression for the first and
second-order DALE, in~\ref{sec:3-2-computational} we explain its
computational benefits, in~\ref{sec:3-3-robustness} its robustness to
OOD sampling and, finally in~\ref{sec:3-4-std}, we quantify the
standard error of the DALE estimation.

\subsection{Differential Aggregate Local Effect (DALE)}
\label{sec:3-1-DALE}
\input{./chapters/3-1-DALE.tex}

\subsection{Computational Benefit}
\label{sec:3-2-computational}
\input{./chapters/3-2-computational.tex}

\subsection{Robustness to out-of-distribution sampling}
\label{sec:3-3-robustness}
\input{./chapters/3-3-robustness.tex}

\subsection{Bias and variance of the DALE approximation}
\label{sec:3-4-std}
\input{./chapters/3-4-std.tex}

% \subsection{Limitations of DALE}
% \label{sec:3-5-limitations}
% \input{./chapters/3-5-limitations.tex}

\section{Experiments}

This section provides an experimental evaluation of DALE in
synthetic and real datasets. In the first part, we generate artificial
data points using a known generating distribution \(p(\mathcal{X}) \)
and a mapping \( f: \mathbf{X} \rightarrow Y \). The second part tests
our proposed method in a real dataset. In this case, we do not possess
ground truth information; threfore, we discuss the qualitative
differences between the two approximations. Code to reproduce the
experiments of this section will become publicly available upon
acceptance.

\subsection{Artificial Examples}
\input{./chapters/4-1-artificial-experiments.tex}

\subsection{The bike-sharing dataset}
\input{./chapters/4-2-real-datasets.tex}

\section{Conclusion}
\input{./chapters/5-conclusion.tex}

\subsubsection*{Acknowledgements}

Redacted.

%% The file named.bst is a bibliography style file for BibTeX 0.99c
\bibliographystyle{named}
\bibliography{ijcai22}

\end{document}
