%\documentclass[wcp,gray]{jmlr} % test grayscale version
\documentclass[wcp]{jmlr}

% givasile packages
\usepackage{bbm}
\usepackage{multirow}
\usepackage{xfrac}
\usepackage[T1]{fontenc}

\usepackage{enumitem}
\usepackage{epstopdf}

\usepackage{bm}
\usepackage{csquotes}
\usepackage{algorithm}
\usepackage{algorithmic}

\newcommand{\dale}{f_{\mathtt{DALE}}}
\newcommand{\ale}{f_{\mathtt{ALE}}}
\newcommand{\alep}{\hat{f}_{\mathtt{ALE}}}
\newcommand{\xc}{\mathbf{x}_c}
\newcommand{\Xcb}{\mathcal{X}_c}
\newcommand{\Xs}{\mathcal{X}_s}
\newcommand{\Xb}{\mathcal{X}}
\newcommand{\xci}{\mathbf{x}^i_{\mathbf{c}}}
\newcommand{\xb}{\mathbf{x}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Jac}{\mathbf{J}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\B}{\mathbb{B}}
\newcommand\todo[1]{\textcolor{red}{#1}}
\usepackage{microtype}

\usepackage{tikz}
\usetikzlibrary{matrix,positioning,arrows.meta,arrows,fit,backgrounds,decorations.pathreplacing}

\tikzset{
  mymat/.style={
    matrix of math nodes,
    text height=2.5ex,
    text depth=0.75ex,
    text width=6.00ex,
    align=center,
    column sep=-\pgflinewidth,
    nodes={minimum height=5.0ex}
  },
  mymats/.style={
    mymat,
    nodes={draw,fill=#1}
  },
  mymat2/.style={
    matrix of math nodes,
    text height=1.0ex,
    text depth=0.0ex,
    minimum width=5ex,
%     text width=7.00ex,
    align=center,
    column sep=-\pgflinewidth
  },
}

\usetikzlibrary{shapes.geometric, arrows, backgrounds, scopes}
\usepackage{pgfplots}
\pgfplotsset{width=6.75cm, compat=newest}
\usepackage[utf8]{inputenc}
\DeclareUnicodeCharacter{2212}{−}
\usepgfplotslibrary{groupplots,dateplot}
\usetikzlibrary{patterns,shapes.arrows}


% The following packages will be automatically loaded:
% amsmath, amssymb, natbib, graphicx, url, algorithm2e

%\usepackage{rotating}% for sideways figures and tables
\usepackage{longtable}% for long tables

% The booktabs package is used by this sample document
% (it provides \toprule, \midrule and \bottomrule).
% Remove the next line if you don't require it.
\usepackage{booktabs}
% The siunitx package is used by this sample document
% to align numbers in a column by their decimal point.
% Remove the next line if you don't require it.
%\usepackage[load-configurations=version-1]{siunitx} % newer version
%\usepackage{siunitx}
%\usepackage{natbib}

% Do not comment the following commands:
\pagenumbering{gobble}
\newcommand{\cs}[1]{\texttt{\char`\\#1}}
\makeatletter
\let\Ginclude@graphics\@org@Ginclude@graphics
\makeatother

\jmlrvolume{}
\jmlryear{2022}
\jmlrworkshop{ACML 2022}

\title[DALE:~Differential Accumulated Local Effects]{DALE:~Differential Accumulated Local Effects for efficient and accurate global model explanations}

 % Use \Name{Author Name} to specify the name.
 % If the surname contains spaces, enclose the surname
 % in braces, e.g. \Name{John {Smith Jones}} similarly
 % if the name has a "von" part, e.g \Name{Jane {de Winter}}.
 % If the first letter in the forenames is a diacritic
 % enclose the diacritic in braces, e.g. \Name{{\'E}louise Smith}

 % Two authors with the same address
 % \author{\Name{Author Name1} \Email{abc@sample.com}\and
 %  \Name{Author Name2} \Email{xyz@sample.com}\\
 %  \addr Address}

 % Three or more authors with the same address:
 % \author{\Name{Author Name1} \Email{an1@sample.com}\\
 %  \Name{Author Name2} \Email{an2@sample.com}\\
 %  \Name{Author Name3} \Email{an3@sample.com}\\
 %  \Name{Author Name4} \Email{an4@sample.com}\\
 %  \Name{Author Name5} \Email{an5@sample.com}\\
 %  \Name{Author Name6} \Email{an6@sample.com}\\
 %  \Name{Author Name7} \Email{an7@sample.com}\\
 %  \Name{Author Name8} \Email{an8@sample.com}\\
 %  \Name{Author Name9} \Email{an9@sample.com}\\
 %  \Name{Author Name10} \Email{an10@sample.com}\\
 %  \Name{Author Name11} \Email{an11@sample.com}\\
 %  \Name{Author Name12} \Email{an12@sample.com}\\
 %  \Name{Author Name13} \Email{an13@sample.com}\\
 %  \Name{Author Name14} \Email{an14@sample.com}\\
 %  \addr Address}


 % Authors with different addresses:
 %  \author{\Name{Author Name1} \Email{abc@sample.com}\\
 %  \addr Address 1
 %  \AND
 %  \Name{Author Name2} \Email{xyz@sample.com}\\
 %  \addr Address 2
 % }

% \editors{Emtiyaz Khan and Mehmet Gonen}

\begin{document}

\maketitle

\begin{abstract}
  Accumulated Local Effect (ALE) is a method for the accurate
  estimation of feature effects, overcoming fundamental failure modes
  of previously-existed methods, such as Partial Dependence Plots
  (PDP). The \textit{approximation} method for ALE, however, faces two
  crucial weaknesses. Firstly, it does not scale well in cases where
  the input has high dimensionality and, secondly, it is vulnerable to
  out-of-distribution sampling in cases with limited training
  examples. In this paper we propose a novel ALE approximation, called
  Differential Accumulated Local Effects (DALE), which overcomes these
  issues. Our proposal has significant computational
  advantages. Specifically, the computational complexity for computing
  the feature effect for all attributes using DALE is similar to what
  ALE requires for a single attribute. DALE, therefore, makes feature
  effect applicable to high-dimensional Machine Learning scenarios
  with near-zero computational overhead. Furthermore, unlike ALE, it
  calculates the feature effect using only the training examples
  resolving misleading estimations due to out-of-distribution
  sampling. Experiments using both synthetic and real datasets
  demonstrate the value of the proposed approach. Code to reproduce
  these experiments is provided in the submission and will become
  publicly available upon acceptance.
\end{abstract}
\begin{keywords}
Feature Effect; Explainable AI; Differentiable Models; Neural-Networks
\end{keywords}

\section{Introduction}
\input{./chapters/1-introduction.tex}
\label{sec:1-introduction}

\section{Related Work}
\input{./chapters/2-related.tex}
\label{sec:2-related}

\section{Background}
\input{./chapters/3-background.tex}
\label{sec:3-feature-effect}

\section{Differential Accumulated Local Effects (DALE)}

In this section, we present DALE.~First, we formulate the expression
for the first and second-order DALE and, then, we explain its
computational benefits and its robustness to OOD sampling. Finally, we
quantify the standard error of the DALE estimation.

\subsection{Method presentation}
\label{sec:4-1-DALE}
\input{./chapters/4-1-DALE.tex}

\subsection{Computational Benefit}
\label{sec:4-2-computational}
\input{./chapters/4-2-computational.tex}

\subsection{Robustness to out-of-distribution sampling}
\label{sec:4-3-robustness}
\input{./chapters/4-3-robustness.tex}

\subsection{Bias and variance}
\label{sec:4-4-std}
\input{./chapters/4-4-std.tex}

% \subsection{Limitations of DALE}
% \label{sec:3-5-limitations}
% \input{./chapters/3-5-limitations.tex}

\section{Experiments}

This section presents the experimental evaluation of DALE using two
synthetic and one real dataset. The experiments aim to compare DALE
(\(f_{\mathtt{DALE}}\)) with ALE approximation
(\(\hat{f}_{\mathtt{ALE}}\)) from the perspectives of both efficiency
and accuracy.

The first synthetic example (Case 1) is designed to compare the
approximations in terms of efficiency. For this reason, we generate
synthetic datasets of varying dimensionality to illustrate the
substantial computational benefits of DALE. The efficiency is
evaluated by measuring the execution time.

The second synthetic example (Case 2) is designed to compare the
approximations in terms of accuracy. More specifically, the example
shows that DALE succeeds in estimating the correct effect when ALE
approximation fails due to OOD sampling. Firstly, we compute the
ground-truth ALE (\(f_{\mathtt{ALE}}\)) directly from
Eq.~\eqref{eq:ALE}\footnote{This is feasible in the synthetic example,
  since we generate artificial data points from a known generating
  distribution \(p(\mathcal{X}) \) and we define a mapping
  \( f: \xb \rightarrow y \) which has a known closed-form.} and,
then, we evaluate the divergence of the two approximations from the
ground-truth using the Mean Squared Error, i.e.,
\(\text{MSE}_{\mathtt{<approx>}} = \E[(\ale -
f_{\mathtt{<approx>}})^2] \), and its normalized version, i.e.,
\(\text{NMSE}_{\mathtt{<approx>}} = \frac{\E[(\ale -
  f_{\mathtt{<approx>}})^2]}{\text{Var}[\ale]}\).

In the last part, we choose the Bike-Sharing dataset to compare DALE
and ALE approximation. We choose this dataset for two
reasons. Firstly, it is the dataset utilised in the original ALE
paper, so it is a proper choice for unbiased comparisons. Secondly, it
contains enough training points to approximate the feature effect
accurately. Therefore, we want to check that \(f_{\mathtt{DALE}}\) and
\(\hat{f}_{\mathtt{ALE}}\) provide similar effects using dense
bins. We also experiment with how both methods behave when using fewer
samples and larger bins. Code to reproduce the experiments is provided
in the supporting material, and it will become publicly available upon
acceptance of the paper.

\subsection{Synthetic Datasets}
\input{./chapters/5-1-artificial-experiments.tex}
\label{sec:5-1-artificial-experiments}

\subsection{Real dataset}
\input{./chapters/5-2-real-datasets.tex}
\label{sec:5-2-real-datasets}

\section{Conclusion and Future Work}
\input{./chapters/6-conclusion.tex}


% \acks{Acknowledgements should go at the end, before appendices and
% references. You can uncomment this for the camera-ready version on
% paper acceptance.}

%\bibliographystyle{plain}
\bibliography{bibliography.bib}

% \appendix
\end{document}