This paper introduced DALE, an efficient and robust-to-OOD
approximation for ALE, the state-of-the-art method for feature effect
analysis. First, we explored the advantages of ALE over the other two
renowned feature effect methods, PDP and MPlots. However, ALE's
approximation scales poorly in big and high-dimensional datasets and
suffers from OOD sampling in cases with limited samples. For
addressing these deficiencies, we proposed DALE, a fast and
on-distribution alternative. We presented the method and discussed the
advantages over the typical ALE approximation. We proved that under
some hypotheses, our proposal is an unbiased estimator of ALE and we
presented a method for quantifying the uncertainty of the explanation,
i.e. the standard error of the approximation. The experiments verify
the aforementioned claims. DALE significantly improves the efficiency
of ALE's approximation by orders of magnitude and secures that local
effect estimations come from on-distribution samples. The latter leads
to more accurate feature effect plots when the bins are wide and the
black-box function changes away from the data generating distribution.

The computational efficiency of DALE delivers a substantial margin for
future extensions. A significant advantage of our proposal is that
effects are computed once on the training set points and can be reused
in different-size bins. The decision for the bin density, i.e. the
resolution of the plot, can be taken afterwards. Therefore, DALE
permits creating feature effect plots at different resolutions with
near-zero computational overhead, which can be embedded into a
multiresolution feature effect plots framework.

In conclusion, we propose DALE as a fast and accurate alternative for
estimating feature effect in cases where the ML model is
differentiable and an auto-differentiable framework is accessible.